---
title: "sgmean: Trimmed Mean Compatible with Statgraphics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sgmean: Trimmed Mean Compatible with Statgraphics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `sgmean` package implements a trimmed mean method that replicates 
the behavior of 'Statgraphics' software. This method differs 
fundamentally from R's built-in `mean(..., trim)` in how it handles 
boundary values.

## The Mathematical Difference

Given a sorted vector of $n$ values and a trim fraction $\alpha$, 
both methods compute $k = \alpha \times n$. The key difference is:

- **R base**: truncates $k$ to the nearest integer and **completely 
removes** that many values from each end.
- **sgmean**: applies a **proportional discount** of $k$ to the 
boundary values, without removing them entirely.

### Formula

$$\bar{x}_{sg} = \frac{x_1(1-k) + x_2 + \cdots + x_{n-1} + x_n(1-k)}{n - 2k}$$

## Practical Example

```{r example}
library(sgmean)

x <- c(10, 20, 30, 40, 50, 60, 70, 80, 90, 200)

# R base method
result_r <- mean(x, trim = 0.05)

# sgmean method (Statgraphics compatible)
result_sg <- sgmean(x, trim = 0.05)

cat("R base  (trim=0.05):", result_r,  "\n")
cat("sgmean  (trim=0.05):", result_sg, "\n")
cat("Difference:         ", abs(result_r - result_sg), "\n")
```

## Why the Results Differ

With `trim = 0.05` and `n = 10`, we get $k = 0.05 \times 10 = 0.5$.

**R base** rounds $k = 0.5$ down to **0** and removes no values:

$$\bar{x}_R = \frac{10+20+30+40+50+60+70+80+90+200}{10} = 65$$

**sgmean** applies a proportional discount of 50% to the boundary values:

$$x_1^* = 10 \times (1 - 0.5) = 5$$
$$x_{10}^* = 200 \times (1 - 0.5) = 100$$
$$\bar{x}_{sg} = \frac{5+20+30+40+50+60+70+80+90+100}{9} = 60.56$$

## When Do Both Methods Agree?

Both methods produce identical results when $k = \alpha \times n$ 
is an exact integer:

```{r comparison}
x <- c(10, 20, 30, 40, 50, 60, 70, 80, 90, 200)

# When k is an exact integer (trim=0.1, n=10, k=1)
cat("trim=0.10 | R base:", mean(x, trim=0.10), 
    "| sgmean:", sgmean(x, trim=0.10), "\n")

# When k is not an integer (trim=0.05, n=10, k=0.5)
cat("trim=0.05 | R base:", mean(x, trim=0.05), 
    "| sgmean:", sgmean(x, trim=0.05), "\n")

# When k is not an integer (trim=0.15, n=10, k=1.5)
cat("trim=0.15 | R base:", mean(x, trim=0.15), 
    "| sgmean:", sgmean(x, trim=0.15), "\n")
```

## Conclusion

The `sgmean` method provides a **continuous and mathematically 
consistent** trimmed mean for any trim fraction between 0 and 0.99, 
replicating the behavior of 'Statgraphics' software and avoiding 
the discontinuities introduced by integer truncation in R's 
built-in `mean()`.
